---
title: "Project 1 Code"
author: "Helene Barrera"
date: "10/1/2020"
output:
  html_document: default
  pdf_document: default
Version: R version 4.0.3
Updated: R version updated 10/10/2020
---

```{r}
#read in data
beer = read.csv("D:/MS in Data Science/Datasets/Beers.csv", header = T)
brew = read.csv("D:/MS in Data Science/Datasets/Breweries.csv", header = T)

#load needed libraries
library(plyr)
library(tidyverse)
library(dplyr)
library(formattable)
library(mice)
library(lattice)
library(pan)
library(ggplot2)
library(scales)
library(ggthemes)
library(knitr)
library(stringr)
library(class)
library(caret)
library(e1071)
library(janitor)
library(highcharter)
library(leaflet)

#convert data to data frame
beer = as.data.frame(beer)
brew = as.data.frame(brew)

#look at how many na values are in the ABV and IBU columns
beer_na_ABV <- sum(is.na(beer$ABV))
beer_na_IBU <- sum(is.na(beer$IBU))
beer_na_ABV
beer_na_IBU

beer$Style <- as.factor(beer$Style)
brew$State <- as.factor(brew$State) 
levels(beer$Style)
str(brew)
```


#1.How many breweries are present in each state?
```{r}
##Take a count of breweries per state, change table headers and format into a nicer looking table
brew_st_table = as.data.frame(table(brew$State))
colnames(brew_st_table) <- c("State", "Frequency")
formattable(brew_st_table)
write.csv(brew_st_table, file = "D:/MS in Data Science/Doing Data Science/Project 1/Question 1 Chart.csv")

```


#2.Merge beer data with the breweries data.
```{r}
##change column name of key beer df to match key of brew df, then join dfs. Edit new col names for clarity.
colnames(beer)[5] <- "Brew_ID"
final <- inner_join(beer,brew,by = "Brew_ID")
colnames(final)[1] <- "Beer_Name"
colnames(final)[8] <- "Brewery_Name"
final <- as.data.frame(final)
final$Ounces <- as.factor(final$Ounces)

##Print the first and last 6 rows

final_first = head(final,6)
final_last = tail(final,6)
first_last = rbind(final_first,final_last) 
formattable(first_last)
```


#3.Address the missing values in each column.
```{r}
##look at the missing variable
ini <- mice(final, maxit = 0)
##this shows 8 complete variables, 1 missing 62 values and 1005 values
table(ini$nmis) 

##Decide to impute the data
##show which variables will be predictors for each missing variable
inlist <- c("ABV","IBU", "Style")
##run prediction matrix
pred <- quickpred(final, minpuc = 0.5, include = inlist)
##Leave out certain predictors
pred[, c("Beer_Name")]=0
pred[, c("Beer_ID")]=0
pred[, c("Brewery_Name")]=0
pred
##run the imputation
imp <- mice(final, maxit = 5, predictorMatrix = pred, seed = 1234)
##merge          
imp_final <- complete(imp)           
##make sure there's no missing values
final_na_ABV <- sum(is.na(imp_final$ABV))
final_na_IBU <- sum(is.na(imp_final$IBU))
final_na_ABV
final_na_IBU
##Found later in the analysis that Scotty K NA from Uncommon Brewers is marketed as a "Low Alcohol Beer" and has a very low ABV. 
##This is removed from the unimputed data since it doesn't have the IBU value. 
##It would be best to remove it from the imputed data as well, since it is it's own category,
##it is a huge outlier that will skew the data, and it is not relevant to what we're looking at.
which.min(imp_final$ABV)  
imp_final[2267,]
imp_final <- imp_final[-c(2267),]

##compare original first/last chart to new one
imp_final_first = head(imp_final,6)
imp_final_last = tail(imp_final,6)
imp_first_last = rbind(imp_final_first,imp_final_last) 
formattable(imp_first_last)

##Convert ABV to percent, since that's how most people are used to seeing it.
imp_final$ABV <- imp_final$ABV*100
```

#4.Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
```{r}
med_ABV <- aggregate(imp_final$ABV, by=list(imp_final$State), FUN = median) 
med_ABV <- as.data.frame(med_ABV)
med_ABV$Group.1 <- as.factor(med_ABV$Group.1)

##Plot it.
med_ABV %>% ggplot(aes(x=Group.1, y=x, )) + geom_bar(stat = 'identity') + 
  ggtitle("Median Alcohol by Volume (ABV) of Beers by State") + theme_few()+
  ylab("ABV %") + xlab("State") + ylim(0, 10) + coord_flip() + 
  scale_x_discrete(limits = rev(levels(med_ABV$Group.1)))

##Same for IBU  
med_IBU <- aggregate(imp_final$IBU, by=list(imp_final$State), FUN = median)
med_IBU <- as.data.frame(med_IBU)
med_IBU$Group.1 <- as.factor(med_IBU$Group.1)
  
med_IBU %>% ggplot(aes(x=Group.1, y=x)) + geom_bar(stat = 'identity') + theme_few() +
  ggtitle("Median International Bitterness Units (IBU) of Beers by State") + xlab("IBU") + ylab("State") +
  ylim(0, 80) + coord_flip() + scale_x_discrete(limits = rev(levels(med_ABV$Group.1)))
  
##do this for unimputed data also 
final <- na.omit(final)
ui_med_ABV <- aggregate(final$ABV, by=list(final$State), FUN = median) 
ui_med_ABV <- as.data.frame(ui_med_ABV)
ui_med_ABV$Group.1 <- as.factor(ui_med_ABV$Group.1)  

ui_med_IBU <- aggregate(final$IBU, by=list(final$State), FUN = median)
ui_med_IBU <- as.data.frame(ui_med_IBU)
ui_med_IBU$Group.1 <- as.factor(ui_med_IBU$Group.1)
```  
  
#5.Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r}
which.max(med_ABV$x)
which.min(med_IBU$x)
##Returned row 9, which is Delaware.
which.max(med_IBU$x)
##Returned row 9, which is also Delaware.
##Compare to unimputed data. Maine is the max ABV and IBU.
which.max(ui_med_ABV$x)
which.max(ui_med_IBU$x)  
```
 
#6.Comment on the summary statistics and distribution of the ABV variable.   
```{r}
summary(imp_final$ABV)
##Which is the lowest and highest ABV?
which.max(imp_final$ABV)
which.min(imp_final$ABV)
imp_final[2279,]
imp_final[770,]


##Histogram
imp_final %>% ggplot(aes(x = ABV)) + geom_histogram(bins = 20, color = 'black', fill = '#1e7b1e') +
              ggtitle("Histogram of ABV%") + xlab("Distribution of ABV") + theme_few()

##same for unimputed  
final$ABV <- final$ABV*100  
summary(final$ABV)
final %>% ggplot(aes(x = ABV)) + geom_histogram(bins = 20, color = 'black') +
              ggtitle("Unimputed Histogram of ABV%") + xlab("Distribution of ABV") + theme_few()
```

#7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? (Scatterplot)   
```{r}
imp_final %>% ggplot(aes(x = ABV, y = IBU)) + geom_point(color = '#1e7b1e') + geom_smooth(method = lm, formula = y ~ x) +
              ggtitle("ABV % by IBU") + xlab("ABV %") + ylab("IBU") + theme_few()
              
##Compare to unimputed data
final %>% ggplot(aes(x = ABV, y = IBU)) + geom_point() + theme_few() +
          ggtitle("ABV % by IBU (non-imputed)") + xlab("ABV %") + ylab("IBU")
```

#8. Compare IPAs to other types of ale. 
```{r}
##Look at how all factors of Style are entered.
view(levels(imp_final$Style))

##Pull out all data with the word "Ale" or "IPA".
temp_Ales <- imp_final
Ales <- temp_Ales %>% filter(str_detect(Style, "(Ale|IPA)")) %>% droplevels()
view(levels(Ales$Style))

##Sort into only 2 factor levels - IPA and Ale.
IPAvALE <- NA
Ales <- cbind(Ales, IPAvALE)

for (i in 1:nrow(Ales)) 
{
  if (is.na(str_match(Ales[i,6], "IPA")))
  {
    Ales[i,11] <- "Ale"  
  }
  else
  {
    Ales[i,11] <- "IPA"
  }
}
Ales$Style <- as.factor(Ales$IPAvALE)

view(Ales)

#Scatterplot of IPA vs Ale
Ales %>% ggplot(aes(x=ABV, y=IBU, color = IPAvALE)) + geom_point(stat = 'identity') +
          theme_few() + ggtitle("Comparison of IPAs vs Ales") +
          scale_color_manual(values = c('#1e7b1e','#d99617'))

##Standardize scales
std_Ales <- Ales
std_Ales$std_ABV = scale(Ales$ABV)
std_Ales$std_IBU = scale(Ales$IBU)


##Make a KNN loop
set.seed(1)
splitPerc = .7

iterations = 100
numk = 50

masterAcc = matrix(nrow = iterations, ncol = numk)

for(j in 1:iterations)
{
accs = data.frame(accuracy = numeric(50), k = numeric(50))
trainIndices = sample(1:dim(std_Ales)[1], round(splitPerc * dim(std_Ales)[1]))
trainAles = std_Ales[trainIndices,]
testAles = std_Ales[-trainIndices,]
for(i in 1:numk)
{
  classifications = knn(trainAles[,c(12,13)],testAles[,c(12,13)],trainAles$IPAvALE, prob = TRUE, k = i)
  table(classifications,testAles$IPAvALE)
  CM = confusionMatrix(table(classifications,testAles$IPAvALE))
  masterAcc[j,i] = CM$overall[1] 
}
}

MeanAcc = colMeans(masterAcc)
plot(seq(1,numk,1),MeanAcc, type = "l")

which.max(MeanAcc)

##Use k=40 for confusion matrix

classifications = knn(trainAles[,c(12,13)],testAles[,c(12,13)],trainAles$IPAvALE, prob = TRUE, k = 40)
  table(classifications,testAles$IPAvALE)
  CM = confusionMatrix(table(classifications,testAles$IPAvALE))
  CM
```
#9. One other useful inference
```{r}

state_brews <- tabyl(Ales, State, IPAvALE)
state_brews$Percent = 100*(state_brews$Ale/(state_brews$Ale + state_brews$IPA))
state_brews <- as.data.frame(state_brews)
state_brews$Percent <- round(state_brews$Percent)

IPA <- Ales %>% filter(IPAvALE == "IPA")
Ale <- Ales %>% filter(IPAvALE == "Ale")
summary(IPA$ABV)
summary(IPA$IBU)
summary(Ale$ABV)
summary(Ale$IBU)
```